---
title: "Lecture 5 - Plotting in R"
output: html_notebook
---

# Topics for Today!
1. Topic: Loading in data
2. Topic: Managing data in dataframes
3. Topic: Basic plotting in R - hist() and boxplot()
4. Topic: Basic plotting in R- A first look at plot()

## Recommended further reading
1. Reading and writing data in R (Ch. 3.1 in Book 1, Ch. 2.6 in Book 2)
2. Managing data from dataframes (Ch. 3.2.4.5+3.2.4.6 in Book 1)
3. Plotting catagorical data with hist(), boxplot() (4.4.2-4 in Book 1, 3.4.1-2 in Book 2)
4. Plotting continuous data with plot() (4.4.1 in Book 1, 4.2.1 in Book 2)

# 1. Topic: Loading Data
```{r}
getwd() # this is the directory in which you can access files
setwd("/Users/mike/Desktop/R Materials/mih140/Lecture 5 - Introduction to Plotting in R/") # this is my directory where I save things.

# Set your directory to whereever you downloaded the RegionEx_Data.txt
# To load the data use the read.table() command
?read.table()
flight_data = read.table("RegionEx_Data.txt", sep = "\t")
flight_data = read.table("RegionEx_Data.txt", header = T, sep = "\t")
# Notice we used the sep = "\t" since the file is a tab separated
# If the file is a .csv, use sep = "," or read.csv(filename)

# To view the table try the View(df) command
View(flight_data)

# Note there are two airlines in the dataset
unique(flight_data$Airline) # this is the unique() function, it returns the unique elements in a vector
```

# 2. Topic: Managing data from dataframes.
```{r}
# Using names() we can get all column names of the flightdata dataset
names(flight_data)

# Using class() we can get the type of the data
class(flight_data$Arrival.delay.in.minutes) #integer
class(flight_data$Airline)                  #factor

# Note the last catagory is a "factor". Factors a special data types for catagorical variables. We can convert other character vectors to factors using the as.factor() method.

flight_data$Delay.indicator = as.factor(flight_data$Delay.indicator)
flight_data$Day.of.Week = as.factor(flight_data$Day.of.Week)

# Lets isolate all flight data for the two airlines

flight_data_MDA = flight_data[flight_data$Airline == "MDA",]
flight_data_RegionEx = flight_data[flight_data$Airline == "RegionEx",]


# We can use the table() function to examine 1 dimension data.
table(flight_data_RegionEx$Delay.indicator)
table(flight_data_RegionEx$Delay.indicator)/nrow(flight_data_RegionEx)
```

# 3. Topic: Basic plotting in R - hist() and boxplot()
```{r}
# Lets reload our data file from before, and isolate the regionEx data.
flight_data = read.table("RegionEx_Data.txt", header = T, sep = "\t", stringsAsFactors = F)
flight_data_RegionEx = flight_data[flight_data$Airline == "RegionEx",]

# To explore the data lets try making a histogram of the delays
delays= flight_data_RegionEx$Arrival.delay.in.minutes

# To make a histogram in R we use the hist() function
# Example. 
hist(delays)

# This makes a very basic histogram. We customize using many of R features.
# We can change the title using the main = "" parameter
# Ex. 
hist(delays, main = "Delays in Minutes")

# Similarly we can change the x and y labels using parameters xlab = "" and ylab = ""
# Ex. 
hist(delays, main = "Delays in Minutes", xlab = "Arrival Delay (min)", ylab = "# of delays")

# Further we can manually set the bins using the parameter breaks = c(b1, b2, .., bk)
# Ex.
hist(delays, main = "Delays in Minutes", 
     xlab = "Arrival Delay (min)", ylab = "# of delays",
     breaks = seq(min(delays)-10,max(delays)+10,10))

# Notice I create bins using seq(min(delays)-10,max(delays)+10,10), run this code to see what it does. Why do I have the -10, +10 in the code?

# Similar to a histogram we can make a box plot using the command boxplot()

# Ex.
boxplot(delays)

# boxplot can be customized just like hist()

# Ex. 
boxplot(delays, main = "Delays in Minutes", 
     xlab = "Arrival Delay (min)", ylab = "# of delays")
```

# 4. Topic: Basic plotting in R - plot()
```{r}
# Plotting continous varibles using plot()
delays= flight_data_RegionEx$Arrival.delay.in.minutes
num_of_passengers = flight_data_RegionEx$Number.of.passengers

plot(delays ~ num_of_passengers) # this plots number of pass on the x vs delays on the y

# We can title it similar to before
plot(delays ~ num_of_passengers, main = "Delay vs. Passenger Load", 
     xlab = "Number of Passengers", ylab = "Delay (min)")

# More on this next time!
```

