---
title: "Lecture 4 - Introduction to R III"
output: html_notebook
---

# Topics for today!

We will demo the final two elements of programming
1. If else statements (Ch. 3.2.1 in Book 1)
2. for loops (Ch.3.2.2, 3.2.4.1 in Book 1)


# 1. Topic: If else statements
```{r}
## Syntax: 
#if(boolean){ 
#  run code here 
#}else{ 
#    run code here
#}
```

## Example 1: Write an if statement to check if variable is a certain size.
```{r}
x = rnorm(1,1,5) # x is one sample from a normal distribution with mean 1 and std 10

if(x >= 1){
  print("x is huge!")
}else{
  print("x is hopelessly tiny")
}
```

If theres more than two cases we can use the else if to chain together commands!
```{r}
## Syntax: 
#if(boolean){ 
#   run code here 
#}else if{ 
#   run code here
#}else{
#   run code here
#}
```


## Example 2: Write an if/else if/else statement to check the size of a variable.
```{r}
# Using x from the previous block!
if(x > 2){
  print("x is huge!")
}else if(x < 0){
  print("x is hopelessly tiny")
}else{
  print("x is boring")
}

# If statements are particular helpful in conjuction with loops and functions.
```

# 2. Topic: for loops
General info: A for loop consists of a *loop varible* (usually we call it i,j, or k) and a *loop vector* (usually 1:N where N is some integer, often).

```{r}
## Syntax: 
# for(loop_var in loop_vec){ 
#   run code 
#   }
```

## Example 1: Print the elements of a vector.
```{r}
data = c(1,1,2,3,5,7)
for(i in data){
  print(i)
}
# In the above loop "i" is the name of the loop variable, and it is pushed through the loop vector named "data" element by element. 
## Loop Trace
# At iteration 1: i = data[1] which is 1, then print(i) is run
# At iteration 2: i = data[2] which is 1, then print(i) is run
# At iteration 3: i = data[3] which is 2, then print(i) is run
# ...
# At iteration 6: i = data[6] which is 7, then print(i) is run and the loop terminates.
```

## Example 2: Remove thhe negative elements from an array. This example will combine if statements with for loops!
```{r}
data = rnorm(10, 0, 1) # 10 samples from a normal random variable
clean_data = rep(0,10)
for(i in 1:length(data)){
  if(data[i] < 0){
    clean_data[i] = NA
  }else{
    clean_data[i] = 1
  }
}
print(clean_data)
```

## Some finer details of for loops
 Recall the motivation for for-loops is to repeat pieces of code in a unified way. In this example we will write a piece of code as a for loop, and then equivalently as a number of code indepent code blocks.

First lets recall some things about strings. Namly, recall from lecture 2 we learned the substring function. Its syntax was substring(char,start_ind, end_ind)

```{r}
strg = "hello"
substring(strg,1,2) # Will return "he"
# Lets learn another R function, nchar(). It will tell us how long a character is (i.e. how many letters are in the string)

# Example: Using nchar()
nchar(strg) # will return 5
```

Now suppose we want to count the number of "l"'s in a given string. For this task we can use for loop to examine all the letters, and everytime we find an "l" we will increment a counter.

## Example 3: Counting "l"'s in a string.
```{r}
string_we_care_about = "lala" # here I'm making up a string, but it will be good because it has l's in it.

N = nchar(string_we_care_about) # N is now the number of letters in the char string_we_care_about

count = 0  # This is a initialization of a counter. Note it is define outside the for loop. If we defined it inside the for loop, it would be reset to 0 every time the loop ran this line.

vec = 1:N # this a vector 1,2,3,..,N. We loop "over" this vector.

for(i in vec){
  if(substring(string_we_care_about,i,i) == "l"){
    count = count + 1 # this line adds 1 to the variable count, then redefines count as the sum
  }
}
print(count)

# Two notes:
## Note 1: the variable i defined in line 48 is the loop variable. It exists to keep track of how many times lines 49-51 are run. To ill
```

Finally to cement these concepts lets trace the execution of the previous example iteration by iteration.

## Example 4: Tracing a for loop
```{r}
string_we_care_about = "lala"
count = 0 # Intialization

## Iteration 1 ##

i = vec[1] # in the first iteration i is equal 1
# Then the following if statement is checked.
if(substring(string_we_care_about,i,i) == "l"){
  # Note substring(string_we_care_about,1,1) = "l", so it will return true
  count = count + 1 # this line adds 1 to the variable count, then redefines count as the sum
}

# At the the end of iteration 1, count = 1, i = 1

## Iteration 2

i = vec[2] # in the second iteration i is equal 2
# Then the following if statement is checked.
if(substring(string_we_care_about,i,i) == "l"){
  # Note substring(string_we_care_about,2,2) = "a", so it will return false
  count = count + 1 # This line thus does not get run
}

# At the the end of iteration 2, count = 1, i = 2

## Iteration 3

i = vec[3] # in the second iteration i is equal 3
# Then the following if statement is checked.
if(substring(string_we_care_about,i,i) == "l"){
  # Note substring(string_we_care_about,3,3) = "l", so it will return true
  count = count + 1 # This line gets run and count is set to 2
}

# At the the end of iteration 3, count = 2, i = 3

## Iteration 4

i = vec[4] # in the second iteration i is equal 4
# Then the following if statement is checked.
if(substring(string_we_care_about,i,i) == "l"){
  # Note substring(string_we_care_about,4,4) = "a", so it will return flase
  count = count + 1 # This line does not get run
}

# At the the end of iteration 4, count = 2, i = 4
# This is the last element of the vector vec, thus the loop ends.

print(count) # at the end of the last iteration, count was equal to 2. Thus 2 is printed.
```

