---
title: "Lecture 9 - Simple Linear Regression"
output: html_notebook
---

# Topics for today!
1. Topic: Checking correlation
2. Topic: Simple Linear Regression Basics
3. Topic: Checking the assumptions of linear regression
4. Topic: Plotting simple linear regression
5. Topic: Linear Regression Example in R


## Data
```{r}
setwd("~/Desktop/R Materials/mih140/Lecture 9 - Linear Regression I")
house = read.table("house.txt", sep = "\t", header = T)

# quick pruning of outliers
house = subset(house, price >= (mean(price) - 3*sd(price)) & price <= (mean(price) + 3*sd(price))) 
```

## 1. Topic: Checking correlation
```{r}
# Test if correlation between price and footage is nonzero.
cor.test(house$price, house$sqft_living)
cor(house[,c("price", "sqft_living")]) # Compute the correlation itself
```

### Note correlation being zero does not imply no relationship, just no linear relationship
```{r}
x = cos(seq(0, 2*pi, .1))
y = sin(seq(0, 2*pi, .1))
plot(x,y)
cor(x,y) # Basically 0, although clearly related!
```


# 2. Topic: Simple Linear Regression Basics
```{r}
# Lets use lin reg to see the relationship between price and living space

reg_model = lm(data = house, price ~ sqft_living) # Simple linear regression
reg_model_2 = lm(data = house, price ~ 0 + sqft_living) # Simple linear regression through the origin

# Examining the regression
reg_model$coefficients # gets the coeff. of the model
reg_model$fitted.values # what the model predicts for each point
avg_error = mean(abs(reg_model$fitted.values - house$price)) # same as mean(abs(reg_model$residuals))
summary(reg_model) # Can read off if coeff. are stat. sign. diff. from 0
```

## 3. Topic: Checking the assumptions of linear regression
```{r}
# The assumptions are:
## 1. Linearity - Look at residuals
## 2. Normality - qqplot of residuals
## 3. Independence (of error term and response) - Next time
## 4. Constant Variance - Look at residuals

# Check assumptions by examining
par(mfrow = c(2,2))
plot(reg_model)

# If the assumptions are satisfied, we can do inference using our model. More next time!
```

## 4. Topic: Plotting simple linear regression
```{r}
# Plotting code from before
price_lim = c(min(house$price), max(house$price))
sqrtft_lim = c(min(house$sqft_living), max(house$sqft_living))
plot(data = house[house$bedrooms <= 3,], price ~ sqft_living, col = "red", xlim = sqrtft_lim, ylim = price_lim)
par(new = T) # tells R to hold prev plot
plot(data = house[house$bedrooms > 3,], price ~ sqft_living, col = "blue", xlim = sqrtft_lim, ylim = price_lim)

# Adding lines to plots
abline(reg_model, col = "black", lwd = 2) # Adds a regression line
abline(0,200, lwd = 2, col = "green") # Adds a second regression line
abline(reg_model_2, lwd = 2, col = "black") # Adds a third regression line
```

## Topic: Linear Regression Example in R
```{r}
setwd("~/Desktop/R Materials/mih140/Lecture 10 - Linear Regression I")
cba = read.table("cba_admissions_1999.txt", sep = "\t", header = T, quote = "", allowEscapes = T)
cba = cba[!is.na(cba$SAT_math),] # remove na rows
```

```{r}
# In this example we examine the relationship betwee sat_math score and sat_verbal score in the cba dataset.

# 1: Fit a simple regression between SAT_verbal and SAT_math
res = lm(cba$SAT_verbal ~ cba$SAT_math)

# 2: Write out the regression equation
summary(res) # observe model Y = 0.46625 X + 290.08827

# 3: Plot the regression line against the scatter plot, be sure to appropriately label your plot!

plot(cba$SAT_math, cba$SAT_verbal, main = "Linear Relationship Between SAT Math and SAT Verbal", xlab = "SAT Math Score", ylab = "SAT Verbal Score")
abline(res, col = "red")

# 4: Plot the residuals, qqplot, comment on trends you see in the data and check assumptions
par(mfrow = c(2,2))
plot(res)
# Residuals seem relatively consistent, the equal variance assumption and independence between our independent variable and the normal noise assumption appears reasonable.
# QQ Plot is straight and diagonal supporting our normality assumption
```

